Controlled Experiments for Word Embeddings

### 2015-10-01 article/cooccurrence-noise-experiment.tex
integration of the corrected noise experiment into article text

\texttt{the domestic \textcolor{red}{CAT\_3} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_2} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_3} is one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization\newline a purebred \textcolor{red}{CAT\_6} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_4} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_4} is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}

>>>>

\texttt{the domestic \textcolor{red}{CAT\_2} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_3} a mostly outdoor \textcolor{red}{CAT\_4} is not \textcolor{red}{CAT\_2} owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_4} is one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization\newline \textcolor{red}{CAT\_6} a purebred \textcolor{red}{CAT\_3} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very \textcolor{red}{CAT\_4} large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_4} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_2} is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively \textcolor{red}{CAT\_5} bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline \textcolor{red}{CAT\_7} in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is \textcolor{red}{CAT\_6} the time at \textcolor{red}{CAT\_3} which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}

### 2015-10-01 article/main.tex
integration of the corrected noise experiment into article text

% compile using latex (xelatex seems to crop the figure strangely)
\documentclass{article} % For LaTeX2e
\usepackage[dvips]{graphicx}
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{mdframed}
\usepackage{rotating}
\usepackage[outdir=./]{epstopdf}

@@

\begin{abstract}
	An experimental approach to studying the properties of word
        embeddings is proposed.  Controlled experiments, achieved
        through modifications of the training corpus, permit the
        demonstration of direct relations between word properties and
        word vector direction and length.  The new approach is
        demonstrated using the word2vec CBOW model with experiments that
        independently vary word frequency and word co-occurrence noise.
        The experiments reveal that word vector length depends more or
        less linearly on both word frequency and the level of noise in
        the co-occurrence distribution of the word.  The coefficients of
        linearity depend upon the word.
\end{abstract}

@@

We show that the length of the word vectors generated by the CBOW model
depends more or less linearly on both word frequency and level of noise
in the co-occurrence distribution of the word.  In both cases, the
coefficient of linearity depends upon the word.  If the co-occurrence
distribution is fixed, then word vector length increases with word
frequency.  If, on the other hand, word frequency is held constant, then
word vector length decreases as the level of noise in the co-occurrence
distribution of the word is increased.  We show furthermore that the
direction of the word vectors is independent of both word frequency and
co-occurrence noise.

@@

\begin{section}{Related work}\label{related-work}
Our notion of word significance is reminiscent of the concept of
``contextual distinctiveness'' introduced earlier by McDonald and
Shillcock \cite{mcdonald2001contextual}.  They define the contextual
distinctiveness of a target word $w$ as the Kullback-Leibler divergence
between the unconditional word distribution $\p(w')$, and the
co-occurrence distribution $\p(w' | w)$ of the target word $w$.  So
defined, contextual distinctiveness is a measure of the amount of
information provided by word $w$ about its contexts of use.  More
specifically, if a word $w$ tends to appear in a wide variety of
contexts, then its co-occurrence distribution tends to diverge less from
unconditional word distribution, and its contextual distinctiveness is
accordingly small.  If, on the other hand, $w$ typically appears in only
a small number of different contexts, its co-occurrence distribution
tends to diverge more from the unconditional word distribution, and a
larger value of contextual distinctiveness results.  To reduce the
computational burden, the authors considered only $690$ target words,
randomly chosen from equally divided frequency bands.  Furthermore, the
authors considered only co-occurrence with the $500$ most frequent
content words.  This significant restriction, necessary in their case in
order to reduce the variance of the estimates of their measure of
contextual distinctiveness, does not apply in our study.

Our experimental finding that word vector length decreases with
co-occurrence noise is related to earlier work by Vecchi, Baroni, and
Zamparelli \cite{vecchi-baroni-zamparelli2011}, where a relation between
vector length and the ``semantic deviance'' of an adjective-noun
composite was studied empirically.  In that paper, which is also based
on word co-occurrence statistics, the authors study adjective-noun
composites.  They built a vocabulary from the 8k most frequent nouns and
4k most frequent adjectives in a large general language corpus and added
22k adjective-noun composites.  For each item in the vocabulary, they
recorded the co-occurrences with the top 10k most frequent content words
(nouns, adjectives or verbs), and constructed word embeddings via
singular value decomposition of the co-occurrence matrix
\cite{landauer-dumais1997}.  The authors considered several models for
constructing vectors of unattested adjective-noun composites, the two
simplest being adding and component-wise multiplying the adjective and
noun vectors.  They hypothesized that the length of the vectors thus
constructed can be used to distinguish acceptable and semantically
deviant adjective-noun composites.  Using a few hundred adjective-noun
composites selected by humans for evaluation, they found that deviant
composites have a shorter vector than acceptable ones, in accordance
with their expectation.  In contrast to their work, our approach does
not require extensive human evaluation.

@@

\begin{subsection}{Replacement procedure}\label{replacement-procedure}
In the experiments detailed below, we modify the corpus in a controlled
manner by introducing tokens into the corpus via a replacement procedure
as follows.  Consider a word, say \word{cat}.  For each occurrence of
this word, a sample $i$, $1 \leqslant i \leqslant n$ is drawn from a
truncated geometric distribution, and that occurrence of the word
\word{cat} is replaced with the token \word{CAT\_i}.  In this way, the
word \word{cat} is replaced throughout the corpus by a family of tokens
with varying frequencies but approximately the same co-occurrence
distribution as \word{cat}.  That is, all these tokens are used in
roughly the same contexts as the original word.

The geometric distribution is truncated to limit the number of tokens
inserted into the corpus.  For any choice $0 < p < 1$ and maximum value
$n > 0$, the truncated geometric distribution is given by the
probability density function
\begin{equation} 
\label{distro}
P_{p, n} (i) = \frac{p^{i-1}  (1-p)}{1 - p^n}, \qquad 1
\leqslant i \leqslant n.
\end{equation} 
The factor in the denominator, which tends to unity in the limit $n \to
\infty$, assures proper normalisation.  We have chosen this distribution
because the probabilities decay exponentially base $p$ as a function of
$i$.  Of course, other distributions might equally well have been chosen
for the experiments.
\end{subsection}
\end{section}

@@

\begin{subsection}{Tokens derived from an artificial, meaningless word}\label{WFVEmeaningless}
Whereas the tokens introduced above all replace an existing word that
carries a meaning, we now include for comparison a high-frequency,
meaningless word.  We choose to introduce an artificial, entirely
meaningless word \word{VOID} into the corpus, rather than choose an
existing (stop)word whose meaninglessness is only supposed.  To achieve
this, we intersperse the word uniformly at random throughout the corpus so
that its relative frequency is $0.005$.  The co-occurrence
distribution of \word{VOID} thus coincides with the unconditional word
distribution.  The replacement procedure is then performed for this
word, using the same values for $p$ and $n$ as above.
Figure~\ref{fig:word-frequency-experiment-text-void} shows the effect of
these modifications on a sample text, where a higher relative frequency of $0.05$ is used instead
for illustrative purposes.  Although the probability of inserting token
\word{VOID\_1} is twice that of inserting token \word{VOID\_2} and four
times that of inserting token \word{VOID\_3}, these three tokens all
appear once in this small piece of text.

\begin{figure}
	\begin{mdframed}
	\input{word-frequency-experiment-text-void.tex}
	\end{mdframed}
	\caption{The same example sentences as in
          Figure~\ref{fig:word-frequency-experiment-text-cat} where
          instead of the word \word{cat} now the meaningless word
          \word{VOID} is replaced with tokens \word{VOID\_i}.  For
          illustrative purposes, the meaningless word \word{VOID} was
          here interspersed with a relative frequency of $0.05$.}
	\label{fig:word-frequency-experiment-text-void}
\end{figure}
\end{subsection}

@@

\subsection{Generating noise}
We take the noise distribution to be the (observed) unconditional word
distribution.  Thus noise can be added to the co-occurrence distribution
of a word by simply interspersing occurrences of that word uniformly at
random throughout the corpus.  A word that is consistently used in a
distinctive context in the unmodified corpus thus appears in the
modified corpus also in completely unrelated contexts.  As in
Section~\ref{WFVE}, we choose a small number of words from the
unmodified corpus for this experiment.
Table~\ref{cooccurrence-noise-words} lists the words chosen, along with
their frequencies in the corpus.

For each of these words, the replacement procedure of
Section~\ref{replacement-procedure} is performed using a geometric rate
of decay of $p = \tfrac{5}{6}$, and truncating the distribution at
$n=8$.  For every replacement token (e.g. \word{CAT\_i}), occurrences of
this token are subsequently interspersed uniformly at random throughout
the corpus, such that the frequency of the replacement token is restored
to that of the original word \word{cat}.  For example, if the original
word \word{cat} occurred $1000$ times, then after the replacement
procedure, \word{CAT\_2} occurs approximately $694$ times, so a further
(approximately) $306$ random occurrences of \word{CAT\_2} are
interspersed throughout the corpus.  Thus the word \word{cat} is removed
from the corpus, and a family of tokens \word{CAT\_i}, $1 \leqslant i
\leqslant 8$ is introduced.  These tokens all have the same frequency as
the word \word{cat}, but their co-occurrence distributions, while based
on that of \word{cat}, have an increasing amount of noise.

@@

The proportion of occurrences of the token \word{CAT\_i} that arise by
replacing the word \word{cat} is $P_{5/6, 8}(i)$, and the proportion
that arise from randomly interspersing this token throughout the corpus
is $r_i := 1 - P_{5/6, 8}(i)$.  The particular value of $p$ (large
compared to that of Section~\ref{WFVE}) was chosen so that the
proportion of noise $r_i$ did not increase too quickly with $i$.  These
proportions are the sampling points on our axis of variation, and this
value of $p$ (together with the choice of $n$) assures a reasonable
spread.  Other parameter values (or indeed other distributions) could,
of course, have been used equally well.

\begin{table}
	\input{noise-cooccurrence-experiment-counts.tex}
	\label{fig:co-occurrence-noise-counts}
	\caption{Words chosen for the co-occurrence noise experiment, along with the occurrence counts in the unmodified corpus. }
	\label{cooccurrence-noise-words}
\end{table}

\begin{figure}
	\begin{mdframed}
	\input{cooccurrence-noise-experiment.tex}
	\end{mdframed}
	\caption{Example sentences modified for the co-occurrence noise
          experiment, where the word \word{cat} was chosen for
          replacement.  The tokens were generated using the distribution
          (\ref{distro}) with $p = \tfrac{5}{6}$ and $n=3$.}
	\label{fig:co-occurrence-noise-experiment-text}
\end{figure}

\begin{subsection}{Experimental results}
Figure~\ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity
of pairs of vectors representing some of the tokens used in this
experiment.  The figure clearly demonstrates that vectors representing
tokens derived from the same word all share a common direction.  Thus
co-occurrence noise has no discernible effect on the direction of word
vectors.

The left panel in Figure~\ref{fig:co-occurrence-noise-graph} reveals
that word vector length varies more or less linearly with the proportion
of noise in the co-occurrence distribution of the word.  This figure motivates an
interpretation of vector length, within a sufficiently narrow frequency
band, as a measure of the absence of co-occurrence noise, or put
differently, of the extent to which a word carries the meaning of a
distinctive context.
The right panel of Figure~\ref{fig:co-occurrence-noise-graph} shows that the
syn1neg vector length does not vary with the proportion of co-occurrence noise.

Recall from Section~\ref{WFVE-results} that the word vector for \word{VOID} is independent of the frequency of \word{VOID}, and is non-zero.
Since the noise distribution coincides with the co-occurrence distribution of \word{VOID}, the word vectors for the experiment words must tend to the word vector for \word{VOID} as the proportion of noise in their co-occurrence distributions approaches $1$.
This is not discernible in Figure~\ref{fig:co-occurrence-noise-heatmap} since the word vector for \word{VOID} is relatively close to the origin.

\begin{figure}
	\includegraphics[scale=0.5]{cooccurrence-noise-heatmap}
	\caption{ Heatmap of the cosine similarity of the vectors
          representing some of the tokens used in the co-occurrence
          noise experiment (the words were chosen at random).  The
          largely red blocks demonstrate that the direction of the word
          vectors does not change when noise is added to the
          co-occurrence distribution.  }
	\label{fig:co-occurrence-noise-heatmap}
\end{figure}

>>>>

% compile using latex (xelatex seems to crop the figure strangely)
\documentclass{article} % For LaTeX2e
\usepackage[dvips]{graphicx}
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{mdframed}
\usepackage{rotating}

\usepackage[outdir=./]{epstopdf}

@@

\begin{abstract}
	An experimental approach to studying the properties of word
        embeddings is proposed.  Controlled experiments, achieved
        through modifications of the training corpus, permit the
        demonstration of direct relations between word properties and
        word vector direction and length.  The approach is demonstrated
        using the word2vec CBOW model with experiments that
        independently vary word frequency and word co-occurrence noise.
        The experiments reveal that word vector length depends more or
        less linearly on both word frequency and the level of noise in
        the co-occurrence distribution of the word.  The coefficients of
        linearity depend upon the word.  The special vector in feature
        space, defined by the (artificial) word with pure noise in its
        co-occurrence distribution, is found to be small but non-zero.
\end{abstract}

@@

We show that the length of the word vectors generated by the CBOW model
depends more or less linearly on both word frequency and level of noise
in the co-occurrence distribution of the word.  In both cases, the
coefficient of linearity depends upon the word.  If the co-occurrence
distribution is fixed, then word vector length increases with word
frequency.  If, on the other hand, word frequency is held constant, then
word vector length decreases as the level of noise in the co-occurrence
distribution of the word is increased.  We show furthermore that the
direction of the word vectors is independent of word frequency and only
moderately depends upon co-occurrence noise.  When noise is added to the
co-occurrence distribution of a word, the corresponding vector smoothly
interpolates between the original word vector and a small vector
perpendicular to it that represents a word with pure noise in its
co-occurrence distribution.  Surprisingly, this special vector in the feature
space, obtained by interspersing a token uniformly at random throughout
the corpus with a frequency sufficiently large to sample all contexts, is
non-zero.

@@

\begin{section}{Related work}\label{related-work}
\mar{move to other paper?}
Our notion of word significance is reminiscent of the concept of
``contextual distinctiveness'' introduced earlier by McDonald and
Shillcock \cite{mcdonald2001contextual}.  They define the contextual
distinctiveness of a target word $w$ as the Kullback-Leibler divergence
between the unconditional word distribution $\p(w')$, and the
co-occurrence distribution $\p(w' | w)$ of the target word $w$.  So
defined, contextual distinctiveness is a measure of the amount of
information provided by word $w$ about its contexts of use.  More
specifically, if a word $w$ tends to appear in a wide variety of
contexts, then its co-occurrence distribution tends to diverge less from
unconditional word distribution, and its contextual distinctiveness is
accordingly small.  If, on the other hand, $w$ typically appears in only
a small number of different contexts, its co-occurrence distribution
tends to diverge more from the unconditional word distribution, and a
larger value of contextual distinctiveness results.  To reduce the
computational burden, the authors considered only $690$ target words,
randomly chosen from equally divided frequency bands.  Furthermore, the
authors considered only co-occurrence with the $500$ most frequent
content words.  This significant restriction, necessary in their case in
order to reduce the variance of the estimates of their measure of
contextual distinctiveness, does not apply in our study.

Our experimental finding that word vector length decreases with
co-occurrence noise is related to earlier work by Vecchi, Baroni, and
Zamparelli \cite{vecchi-baroni-zamparelli2011}, where a relation between
vector length and the ``semantic deviance'' of an adjective-noun
composite was studied empirically.  In that paper, which is also based
on word co-occurrence statistics, the authors study adjective-noun
composites.  They built a vocabulary from the 8k most frequent nouns and
4k most frequent adjectives in a large general language corpus and added
22k adjective-noun composites.  For each item in the vocabulary, they
recorded the co-occurrences with the top 10k most frequent content words
(nouns, adjectives or verbs), and constructed word embeddings via
singular value decomposition of the co-occurrence matrix
\cite{landauer-dumais1997}.  The authors considered several models for
constructing vectors of unattested adjective-noun composites, the two
simplest being adding and component-wise multiplying the adjective and
noun vectors.  They hypothesized that the length of the vectors thus
constructed can be used to distinguish acceptable and semantically
deviant adjective-noun composites.  Using a few hundred adjective-noun
composites selected by humans for evaluation, they found that deviant
composites have a shorter vector than acceptable ones, in accordance
with their expectation.  In contrast to their work, our approach does
not require human annotation.

@@

\begin{subsection}{Replacement procedure}\label{replacement-procedure}
In the experiments detailed below, we modify the corpus in a controlled
manner by introducing tokens into the corpus via a replacement
procedure.  For the frequency experiment, the procedure is as follows.
Consider a word, say \word{cat}.  For each occurrence of this word, a
sample $i$, $1 \leqslant i \leqslant n$ is drawn from a truncated
geometric distribution, and that occurrence of the word \word{cat} is
replaced with the token \word{CAT\_i}.  In this way, the word \word{cat}
is replaced throughout the corpus by a family of tokens with varying
frequencies but approximately the same co-occurrence distribution as
\word{cat}.  That is, all these tokens are used in roughly the same
contexts as the original word.

The geometric distribution is truncated to limit the number of tokens
inserted into the corpus.  For any choice $0 < p < 1$ and maximum value
$n > 0$, the truncated geometric distribution is given by the
probability density function
\begin{equation} 
\label{distro}
P_{p, n} (i) = \frac{p^{i-1}  (1-p)}{1 - p^n}, \qquad 1
\leqslant i \leqslant n.
\end{equation} 
The factor in the denominator, which tends to unity in the limit $n \to
\infty$, assures proper normalisation.  We have chosen this distribution
because the probabilities decay exponentially base $p$ as a function of
$i$.  Of course, other distributions might equally well have been chosen
for the experiments.

For the noise experiment, we take instead of a geometric distribution,
the distribution
\begin{equation}
\label{distro2} 
P_n (i) = \frac{2 (n-i)}{n(n-1)} , \qquad 1 \leqslant i \leqslant n.
\end{equation} 
We have chosen this distribution for the noise experiment, because it
leads to evenly spaced proportions of co-occurrence noise that cover the entire interval
$[0,1]$.
\end{subsection}
\end{section}

@@

\begin{subsection}{Tokens derived from an artificial, meaningless word}
\label{WFVEmeaningless}
Whereas the tokens introduced above all replace an existing word that
carries a meaning, we now include for comparison a high-frequency,
meaningless word.  We choose to introduce an artificial, entirely
meaningless word \word{VOID} into the corpus, rather than choose an
existing (stop)word whose meaninglessness is only supposed.  To achieve
this, we intersperse the word uniformly at random throughout the corpus
so that its relative frequency is $0.005$.  The co-occurrence
distribution of \word{VOID} thus coincides with the unconditional word
distribution.  The replacement procedure is then performed for this
word, using the same values for $p$ and $n$ as above.
Figure~\ref{fig:word-frequency-experiment-text-void} shows the effect of
these modifications on a sample text, where a higher relative frequency
of $0.05$ is used instead for illustrative purposes.  Although the probability of inserting token
\word{VOID\_1} is twice that of inserting token \word{VOID\_2} and four
times that of inserting token \word{VOID\_3}, these three tokens all
appear once in this small piece of text.

\begin{figure}
	\begin{mdframed}
	\input{word-frequency-experiment-text-void.tex}
	\end{mdframed}
	\caption{The same example sentences as in
          Figure~\ref{fig:word-frequency-experiment-text-cat} where
          instead of the word \word{cat} now the meaningless word
          \word{VOID} is replaced with tokens \word{VOID\_i}.  For
          illustrative purposes, the meaningless word \word{VOID} was
          here interspersed with a relative frequency of
          $0.05$. }
	\label{fig:word-frequency-experiment-text-void}
\end{figure}
\end{subsection}

@@

\subsection{Generating noise}
We take the noise distribution to be the (observed) unconditional word
distribution.  Noise can then be added to the co-occurrence distribution
of a word by simply interspersing occurrences of that word uniformly at
random throughout the corpus.  A word that is consistently used in a
distinctive context in the unmodified corpus thus appears in the
modified corpus also in completely unrelated contexts.  As in
Section~\ref{WFVE}, we choose a small number of words from the
unmodified corpus for this experiment.
Table~\ref{cooccurrence-noise-words} lists the words chosen, along with
their frequencies in the corpus.

For each of these words, the replacement procedure of
Section~\ref{replacement-procedure} is performed using the
distribution~(\ref{distro2}) with $n=7$.  For every replacement token
(e.g. \word{CAT\_i}), additional occurrences of this token are
interspersed uniformly at random throughout the corpus, such that the
final frequency of the replacement token is $2/n$ times that of the
original word \word{cat}.  For example, if the original word \word{cat}
occurred $1000$ times, then after the replacement procedure,
\word{CAT\_2} occurs approximately $238$ times, so a further
(approximately) $2/7 \times 1000 - 238 \approx 48$ random occurrences of
\word{CAT\_2} are interspersed throughout the corpus.  In this way, the
word \word{cat} is removed from the corpus and replaced with a family of
tokens \word{CAT\_i}, $1 \leqslant i \leqslant 7$.  These tokens all
have the same frequency, but their co-occurrence distributions, while
based on that of \word{cat}, have an increasing amount of noise.
Specifically, the proportion of noise for the $i$th token is
%
$$ 1 - \frac{n}{2} P_n(i) = \frac{i-1}{n-1} , \;\; \textrm{ or } \;\; 0,
\frac{1}{n-1}, \frac{2}{n-1}, \ldots, 1 \;\; \textrm{ for } \;\; i =
1,2, \ldots, n , 
$$ 
%
which is evenly distributed.  The first token contains no noise at all,
while the last token stands for pure noise.  The particular choice of
$n$ assures a reasonable coverage of the interval $[0,1]$.  Other
parameter values (or indeed other distributions) could, of course, have
been used equally well.

@@

\begin{table}
	\input{noise-cooccurrence-experiment-counts.tex}
	\label{fig:co-occurrence-noise-counts}
	\caption{Words chosen for the co-occurrence noise experiment,
          along with the word frequencies in the unmodified corpus. }
	\label{cooccurrence-noise-words}
\end{table}

\begin{figure}
	\begin{mdframed}
	\input{cooccurrence-noise-experiment.tex}
	\end{mdframed}
	\caption{Example sentences modified for the co-occurrence noise
          experiment, where the word \word{cat} was chosen for
          replacement.  The tokens were generated using the distribution
          (\ref{distro2}) with $n=7$.}
	\label{fig:co-occurrence-noise-experiment-text}
\end{figure}

\begin{subsection}{Experimental results}
%
Figure~\ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity
of pairs of vectors representing some of the tokens used in this
experiment.  Remember that the first token ($i=1$) in a family is
without noise in its co-occurrence distribution, while the last one
($i=n$, with $n=7$) stands for pure noise and has therefore no relation
anymore with the word it derives from.  The figure demonstrates that the
vectors within a family only moderately deviate from the original
direction defined by the first token ($i=1$) when noise is added to the
co-occurrence distribution.  For $1<i<7$, the deviation typically
increases with the proportion of noise.  The vector of the last token
($i=n$), associated with pure noise, is seen within each of the families
to point in a completely different direction, more or less perpendicular
to the original one.  To understand this interpolating behavior, recall
from Section~\ref{WFVE-results} that the vector for the entirely
meaningless word \word{VOID} is small but non-zero.  Since the noise
distribution coincides with the co-occurrence distribution of
\word{VOID}, the vectors for the experiment words must tend to the word
vector for \word{VOID} as the proportion of noise in their co-occurrence
distributions approaches $1$.
This convergence to a common point is only indistinctly apparent in
Figure~\ref{fig:co-occurrence-noise-heatmap} as the frequency of the experiment
tokens is insufficient to sample the full variety of the contexts of VOID, i.e.
all contexts (see Section \ref{WFVE-direction}).

The left panel in Figure~\ref{fig:co-occurrence-noise-graph} reveals
that vector length varies more or less linearly with the proportion of
noise in the co-occurrence distribution of the word.  This figure
motivates an interpretation of vector length, within a sufficiently
narrow frequency band, as a measure of the absence of co-occurrence
noise, or put differently, of the extent to which a word carries the
meaning of a distinctive context.

\begin{figure}
	\includegraphics[scale=0.5]{cooccurrence-noise-heatmap}
	\caption{ Heatmap of the cosine similarity of the vectors
          representing some of the tokens used in the co-occurrence
          noise experiment (the words were chosen at random).  The
          largely red blocks demonstrate that for $i<7$ the direction of
          the vectors only moderately changes when noise is added to the
          co-occurrence distribution.  The vector of the tokens
          associated with pure noise ($i=7$) is seen to be almost
          perpendicular to the word vectors they derive from. }
	\label{fig:co-occurrence-noise-heatmap}
\end{figure}

### 2015-10-01 article/word-frequency-experiment-text-cat.tex
integration of the corrected noise experiment into article text

\texttt{the domestic \textcolor{red}{CAT\_1} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_1} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_1} is one whose ancestry is recorded by a \textcolor{red}{CAT\_2} fancier organization\newline a purebred \textcolor{red}{CAT\_1} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_3} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_3} is usually dominant over the others}

>>>>

\texttt{the domestic \textcolor{red}{CAT\_2} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_1} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_1} is one whose ancestry is recorded by a \textcolor{red}{CAT\_2} fancier organization\newline a purebred \textcolor{red}{CAT\_2} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_4} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_1} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} is usually dominant over the others}

### 2015-10-01 article/word-frequency-experiment-text-void.tex
integration of the corrected noise experiment into article text

\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the cat cannot produce taurine\newline \textcolor{red}{VOID\_1} within groups one cat is usually dominant over the others}

>>>>

\texttt{\textcolor{red}{VOID\_1} the domestic cat was first classified as felis catus\newline the semiferal cat \textcolor{red}{VOID\_3} a mostly outdoor cat is not \textcolor{red}{VOID\_2} owned by \textcolor{red}{VOID\_1} any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among \textcolor{red}{VOID\_1} mammals in having very large eye sockets\newline another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others}
